---
title: "Reproducible Research: Peer Assessment 1"
author: "Yalun"
date: "5/7/2020"
output: html_document
keep_md: true
---


## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv", stringsAsFactors = FALSE)

activity.comp <- activity[complete.cases(activity),]
activity.steps <- activity.comp[,1:2]

result.sum <- aggregate(activity.steps$steps,
                    by=list(activity.steps$date),
                    FUN=sum)
result.sum$Group.1 <- as.Date(result.sum$Group.1)
```

Load 'activity' data from the csv file, then process data to show the total number of steps taken each day.

## What is mean total number of steps taken per day?

```{r fig.cap="Histogram/barplot of the total nuber of steps taken each day"}
library(ggplot2)
library(gridExtra)
p1 <- qplot(result.sum$x,geom="histogram",xlab="steps",ylab="frequency",bins=20)
p2 <- ggplot(result.sum, aes(Group.1,x))+geom_col()+ylab("steps")+xlab("date")
grid.arrange(p1,p2, nrow = 2)

```
### Calculate mean and median:

```{r}
summary(result.sum$x)
```
Therefore mean is 10766 and median is 10765 steps taken per day

## What is the average daily activity pattern?
```{r}
result.average <- aggregate(activity.comp$steps,
                        by=list(activity.comp$interval),
                        FUN=mean)
ggplot(result.average, aes(Group.1,x))+geom_line()+ylab("averaged steps")+xlab("interval")
```
```{r}
step.max <- which.max(result.average$x)
result.average[step.max,]
```
Therefore, the '835' interval contains the maximum number of steps.


## Imputing missing values

Total number of missing days:
```{r}
sum(is.na(activity))
```
Filling missing values in the datasheet:

```{r}
library(tidyr)
for (i in 1:nrow(activity)){
        if (is.na(activity[i,1])==TRUE){
                thatinterval=activity[i,3]
                thatmean=result.average[match(thatinterval,result.average$Group.1),2]
                activity[i,1]=replace_na(NA,thatmean)
        }
}
head(activity)
```

Histogram of the filled datasheet:
```{r}
result.sum2 <- aggregate(activity$steps,
                        by=list(activity$date),
                        FUN=sum)

result.sum2$Group.1 <- as.Date(result.sum2$Group.1)

p3 <- qplot(result.sum2$x,geom="histogram",xlab="steps",ylab="frequency",bins=20)
p4 <- ggplot(result.sum2, aes(Group.1,x))+geom_col()+ylab("steps")+xlab("date")
grid.arrange(p3,p4, nrow = 2)
```
The mean and median total number of steps taken per day:
```{r}
summary(result.sum2$x)
```
The filling the datasheet with averaged data barely changes the mean and median.


## Are there differences in activity patterns between weekdays and weekends?
```{r message=FALSE}
library(dplyr)
activity <- mutate(activity,week=weekdays(as.Date(activity$date)))
activity.weekday <- subset(activity, week=="Monday"|week=="Tuesday"|week=="Wednesday"|week=="Thursday"|week=="Friday")
activity.weekend <- subset(activity, week=="Saturday"|week=="Sunday")

weekday.average <- aggregate(activity.weekday$steps,
                            by=list(activity.weekday$interval),
                            FUN=mean)
weekend.average <- aggregate(activity.weekend$steps,
                             by=list(activity.weekend$interval),
                             FUN=mean)

p5<- ggplot(weekday.average, aes(Group.1,x))+geom_line()+ylab("weekday steps")+xlab("interval")
p6<- ggplot(weekend.average, aes(Group.1,x))+geom_line()+ylab("weekend steps")+xlab("interval")
grid.arrange(p5,p6, nrow = 2)
```

There seems to be more steps on weekends than on weekdays
